# Methods

This R Markdown file documents the entire data processing and analysis pipeline for "Population health implications of exposure to pervasive military aircraft noise pollution".

## Dependencies

R software version 4.3.0 was used for this study. Noisemap software is available from the [US Department of Defense Community and Environmental Noise Primer Resources](http://dodnoise.org/primer_resources). BaseOps version 7.368 was used as the graphical user interface for Noisemap simulation data entry and management in concert with a) Omega10 and Omega11 to calculate sound over distance for aircraft flight operations, ground maintenance, and run-up operations; b) NMap 7.3 and MRNMap to calculate noise exposure values on the ground; c) NMPlot 4.974 to convert calculated noise exposure values to noise contour plots.

```{r}
# Install package dependencies
packages = c(
  'dplyr',      # 1.1.2
  'fasterize',  # 1.0.4
  'gdalUtils',  # 2.0.3.2
  'ggplot2',    # 3.4.2
  'ggpmisc',    # 0.5.2
  'ggpp',       # 0.5.2
  'ggrepel',    # 0.9.3
  'glue',       # 1.6.2
  'leafem',     # 0.2.0
  'mapview',    # 2.11.0
  'patchwork',  # 1.1.2
  'pdftools',   # 3.3.3
  'raster',     # 3.6-23
  'readr',      # 2.1.4
  'readxl',     # 1.4.2
  'rslurm',     # 0.6.2
  'scales',     # 1.2.1
  'sf',         # 1.0-13
  'sp',         # 1.6-1
  'stringr',    # 1.5.0
  'tidycensus', # 1.4.1
  'tidyr',      # 1.3.0
  'tigris',     # 2.0.3
  'viridis',    # 0.6.3
  'zoo'         # 1.8-12
)
install.packages(packages[!(packages %in% rownames(installed.packages()))])
```

### Database configuration and data access

First, set the `database_path` global variable in `global.R` to your local database path. This database should include the following directories:

-   `NAVY` - The NAVFAC database, with acoustic data files converted from Larson Davis binary format .LD0 to .xlsx.
-   `NPS` - Acoustic data from the National Park Service (.txt).
-   `JGL` - Acoustic data from JGL Acoustics, Inc (.csv).
-   `GIS` - GIS data, including subdirectories:
    -   `NCLD` - [US National Land Cover Database](https://www.mrlc.gov/data/nlcd-2019-land-cover-conus)
    -   `NCES` - [NCES school location shapefiles](https://nces.ed.gov/programs/edge/geographic/schoollocations)

Statement about how data are formatted (i.e. site-date, 1-second SPL measurements, with frequency spectrum when available). These routines are written...

```{r}
# Configure and check for mounted database
source('global.R')
stopifnot(dir.exists(database_path))

# Display monitoring location map
data_sites = get_data_sites()
mapview(
  data_sites,
  xcol='Longitude', ycol='Latitude', zcol='Org',
  layer.name = 'Organization', crs=4269, grid=F, legend=T,
  col.regions=c('yellow', 'navy', 'green')
) %>% addStaticLabels(label=data_sites$ID, direction='top')

# Create tables mapping raw data files to sites and date
source('data/load/load_file_jgl.R')
file_map_jgl = map_files_jgl_csv()
source('data/load/load_file_navy.R')
file_map_navy = map_files_navy_csv()
source('data/load/load_file_nps.R')
file_map_nps = map_files_nps_csv()

# Load raw data for each site date and save as standardized .csv
source('data/load/load_site_date.R')
create_site_date_csvs('JGL')
create_site_date_csvs('NAVY')
create_site_date_csvs('NPS')
```

## Acoustic monitoring metrics (noise regime)

### Cumulative acoustic metrics

```{r}
# Calculate cumulative metrics for each site date
source('analysis/noise_regime/preprocessing/calculate_cumulative_metrics.R')
calculate_cumulative_metrics_csv('JGL')
calculate_cumulative_metrics_csv('NAVY')
calculate_cumulative_metrics_csv('NPS')

# Evaluate cumulative metrics
source('analysis/noise_regime/evaluate_cumulative_metrics.R')
```

### Single-event acoustic metrics

```{r}
# Calculate single-event metrics for each site date
source('analysis/noise_regime/preprocessing/calculate_single_event_metrics.R')
calculate_events_csv('NAVY')
calculate_events_csv('JGL')
calculate_events_csv('NPS')

# Retrieve noise events reported by Navy
source('data/events/load_events_navy.R')
source('data/events/load_events_navy_moa.R')

# Evaluate single-event metrics
source('analysis/noise_regime/evaluate_single_event_metrics.R')
```

## Aircraft operations data and simulation models

### Temporal distribution analysis

```{r}
# Retrieve flight operations reported by Navy
source('data/flight_ops/load_ops.R')

# Evaluate temporal distribution of noise and aircraft operations
source('analysis/noise_regime/temporal_distribution.R')
```

### Aggregate aircraft operations data

1.  Download "[V. Noise Modeling Data - PUBLIC_NOISEMAP](https://www.navfac.navy.mil/Portals/68/Documents/Business-Lines/Asset-Management/Sound/Remaining-Adds/PUBLIC_NOISEMAP.zip?ver=KEbUPIKWwvnjZl0H4vUg9g%3d%3d)" from NAVY database and unzip.
2.  Format flight operations
    1.  Input: NOISEMAP/NASWI/MP[N]/\*Flight Operations.xml, where 'N' is the monitoring period, 1-4
    2.  Output: .csv for each Flight Operations.xml file, located under 'data/simulation/flight_ops'
    3.  For each Flight Operations.xml file, open in Microsoft Excel. If prompted, do not update workbook links. Then, select File \> Save As \> Comma Separated Values (.csv), and save to the corresponding directory in 'data/simulation/flight_ops/MP[N]'
3.  Run `data/simulation/flight_ops/aggregate_flight_ops.R` to generate the per-period aggregates and the total combined average .csv outputs, respectively.
    1.  Input: 'data/simulation/flight_ops/\*.csv'

    2.  Outputs: 'data/simulation/flight_ops/\_output/NASWI_MP[N]\_Noisemap - Flight Operations Aggregated.csv', 'data/simulation/flight_ops/\_output/NASWI_Noisemap - Flight Operations Combined Average.csv', 'data/simulation/flight_ops/\_output/NASWI_Noisemap - Flight Operations Combined Average - Night Only.csv'; and also simulated adjustments (0.5 and 1.5)

```{r}
# Generate aggregated flight operations data
source('data/simulation/aggregate_flight_ops.R')
```

### Noise simulation modeling with Noisemap

1.  Next, open the 'data/simulation/\_output/csv/\* Combined Average \*.csv' files with Microsoft Excel and re-save as .xml files in data/simulation/\_output/xml.
2.  Open 'simulation/baseops/DNL/NASWI_Combined_Average_DNL.baseops' with BaseOps. Note that modeled static operations were constant across all four monitoring periods. To import new flight operations into BaseOps:
    1.  File \> Import Flight Operations from Spreadsheet
        1.  Option Categories \> File \> Import operations from the following spreadsheet file: \<the xml file you just saved, either the total combined average for DNL and LEQ24, or the "Night Only" for LNIGHT\>
        2.  Option Categories \> Columns:
            1.  Flight Profile Name Column: B (2)
            2.  Num Day Ops Column: H (8)
            3.  Num Night Ops Column: I (9)
        3.  Also import flight tracks: yes
            1.  Flight Track Name Column: C (3)
        4.  Option Categories \> Missing Data
            1.  If a flight profile in the spreadsheet is missing from the BaseOps case, then... Add the missing profile to the BaseOps case
            2.  If a flight profile in the BaseOps case is missing from the spreadsheet, then... Leave the profile unchanged in the BaseOps case
            3.  If you are also importing flight tracks, and a flight track in the spreadsheet is missing from the BaseOps case, then... Set the profile's flight track to "undefined"
        5.  Press OK... You should see the following message: `Importing flight profiles from spreadsheet file NASWI_MP1_Noisemap - Flight Operations.xml.  The following flight profiles appear in both the BaseOps case and the spreadsheet file.  The daily flight profile operation counts in the BaseOps case will be updated to match the values in the spreadsheet file.`

#### Run BaseOps simulations and generate NMPlot noise contour maps

*See NoiseMap and AEDT Gap Analysis Technical Report for further details of NoiseMap and BaseOps software. NoiseMap BaseOps cases have been made for DNL and Leq24 (cumulative, see 4.1.1.1) and Lnight (9hr, see 4.1.1.2).*

1.  For a given .baseops file under 'simulation/baseops' (DNL, LEQ24, or LNIGHT), open it with BaseOps, then select `Case > Run Case`
2.  After the case has finished running, select `Case > Plot`. This will open the default contours in NMPlot.
3.  Click the `Edit Options` button. From `Option Categories > Contours > Levels`, manually specify contour levels with a lowest primary level (such as 35) and highest 150, with spacing between primary levels of 1, and 0 secondary levels. Click `Apply`.
    1.  NOTE: To view the contours on a map, you can select Background \> Layers \> Add Layer (SlippyMap) \> Options Title Server - URL Prefix: tile.openstreetmap.org Attribution Text: OpenStreetMap
4.  Select `File > Export to GIS`. Under format, select `ESRI ARC/INFO Shapefile (SHP)`, then click `Properties`
    1.  Coordinate System \> Export in the following coordinate system... WGS 84
    2.  Level of Detail \> Export with a level of detail appropriate for display at a scale of...
        -   1:24000
        -   1 inch = 2000 feet
    3.  Create based upon the filename '....\\simulation\\\_output\\DNL\\DNL' , replacing 'DNL' with the corresponding metric.
5.  Click `OK`

#### Simulation model validation

```{r}
# Validate simulation against results from the Navy’s monitoring study
source('simulation/validation.R')
```

## Population noise exposure

### Dasymetric population density

```{r}
# Determine exposed areas
source('simulation/contours.R')
exposed_areas = get_exposed_areas()

# Generate dasymetric population density maps for all exposed areas
source('analysis/population_noise_exposure/dasymetric_population_density.R')
for (id in exposed_areas$counties) generate_dasypop_county(id)
for (id in exposed_areas$native_lands) generate_dasypop_native_land(id)
```

Combine dasymetric population density maps with noise contour maps to yield noise exposure across the population.

```{r}
source('analysis/population_noise_exposure/calculate_population_noise_exposure.R')
source('analysis/population_noise_exposure/assess_population_noise_exposure.R')
```

## Population health impacts

Next, conduct health assessment for chosen metrics across the population.

```{r}
source('analysis/population_health_impacts/calculate_population_health_impacts.R')
source('analysis/population_health_impacts/assess_population_health_impacts.R')
```

Conduct childhood learning school impact evaluation.

```{r}
source('analysis/population_health_impacts/childhood_learning.R')
```

Conduct OSHA and NIOSH hearing loss impact evaluation.

```{r}
source('analysis/population_health_impacts/preprocessing/calculate_osha_niosh.R')
source('analysis/population_health_impacts/hearing_impairment.R')
```
